---
title: "Chapter-3-Exercises"
author: "Gavin McCorry"
date: "2024-02-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Atomic Vectors
## Example is the die vector
```{r Vectors}
die <- c(1, 2, 3, 4, 5, 6)
die

is.vector(die)
```

## Vectors of length 1
```{r vects length 1}
five <- 5
five

is.vector(five)

length(five)

length(die)
```


## Integer and Character Vectors
```{r Int and Char Vects}
int <- 1L # Make Integeer Vector by adding L
text <- "ace" # make Character vector by surrounding in quotes

int <- c(1L, 5L)
text <- c("ace", "hearts")

sum(int)
# sum(text)
# Error in `sum()`:
# ! invalid 'type' (character) of argument
# Execution halted
```

# Doubles
```{r Doubles}
#EX:
die <- c(1, 2, 3, 4, 5, 6)
die

typeof(die)
```

# Integers
```{r Integers}
int <- c(1L, 2L, 4L)
int
typeof(int)

sqrt(2)^2 - 2
```

# Characters
```{r Char}
text <- c("Hello", "World")
text

typeof(text)

typeof("World")
```


# Logicals
```{r}
3 > 4

logic <- c(TRUE, FALSE, TRUE)
logic

typeof(logic)

typeof(F)
```

# Complex and Raw
```{r}
comp <- c(1 +1i, 1 + 2i, + 3i)
comp

typeof(comp)

raw(3)

typeof(raw(3))
```


# Exercise
## Create an atomic vector that stores just the face names of the cards in a royal flush,
## for exampl, the ace of spades, king of spads, queen of spades, jack of spades, and ten of spades.
## The face of the ace of spades would be "ace", and "spades" is the suit
```{r Exercise 1}
hand <- c("ace", "king", "queen", "jack", "ten")
hand

typeof(hand)
```


# Attributes
## Names - Most common attribute
```{r Names}
names(die)

names(die) <- c("one", "two", "three", "four", "five", "six")

names(die)

attributes(die)

die

die + 1 # names wont effect the actuak value of the vector

# you can rename the elements vector
names(die) <- c("uno", "dos", "tres", "cuatro", "cinco", "seis")
die

# can remove all names as well
names(die) <- NULL
die
```


## Dim
```{r}
# can transform an atomic vector into an n-dimensional array

# 2x3 matrix
dim(die) <- c(2, 3)
die

# 3x2 matrix
dim(die) <- c(3, 2)
die

# 1 x 2 x 3 hyperecube
dim(die) <- c(1, 2, 3)
die
```

## Matrices
```{r}
m <- matrix(die, nrow = 2)
m

m <- matrix(die, nrow = 2, byrow = TRUE)
m
```

## Arrays
```{r}
ar <- array(c(11:14, 21:24, 31:34), dim = c(2, 2, 3))
ar
```

# Exercise 2
```{r Exrcise 2}
hand1 <- c("ace", "king", "queen", "jack", "ten", "spades", "spades", "spades", "spades", "spades")

matrix(hand1, nrow = 5)
```

# Class
```{r}
dim(die) <c(2, 3)
typeof(die)

class(die)

attributes(die)

class("Hello")

class(5)
```

# Dates and Times
```{r}
now <- Sys.time()
now

typeof(now)

class(now)

unclass(now)

# What day is it a million seconds after 12:00 am han 1, 1970?
mil <- 1000000
mil

class(mil) <- c("POSIXct", "POSIXt")
mil
```

# Factors
```{r}
gender <- factor(c("male", "female", "female", "male"))
typeof(gender)

attributes(gender)

unclass(gender)

gender

as.character(gender)
```

# Exercise 3
```{r Exercise 3}
card <- c("ace", "hearts", 1)
card
# you cant put multiple types of data in an atomic vector
```

# Coercion
```{r}
# R uses the same coercion ruls when you try to do math with logic values.So:
sum(c(TRUE, TRUE, FALSE, FALSE))

# is the same as:
sum(c(1, 1, 0, 0))

as.character(1)
as.logical(1)
as.numeric((FALSE))
```


# Lists
```{r}
list1 <- list(100:300, "R", list(TRUE, FALSE))
list1
```

# Exercise 4
```{r Exercise 4}
# make one card using a list
card <- list("ace", "hearts", 1)
card
```

# Data Frames
```{r}
df <- data.frame(fac = c("ace", "two", "six"), suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 6))
df

typeof(df)
class(df)
str(df)

df <- data.frame(fac = c("ace", "two", "six"), suit = c("clubs", "clubs", "clubs"), value = c(1, 2, 6), stringsAsFactors = FALSE)
df
```

# Loading Data
## load deck.csv from http://bit.ly/deck_CSV
```{r}
# head(deck)
```


# Saving Data
```{r}
# write.csv(deck, file = "cards.csv", row.names = FALSE)
```



