---
title: "Chapter 5"
author: "Gavin McCorry"
date: "2024-03-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 5: Modifying Values
### Loading Deck
```{r}
deck <- read.csv("C:/Users/gwmcc/OneDrive/Documents/GitHub/Data-332/Chapter-3/cards.csv")

```
### Shuffle Function from Chapter 4
```{r}
shuffle <- function(cards) {
  random <- sample(c(1:52), size = 52)
  cards[random,]
}
```


## Changing Values In Place
```{r}
vec <- c(0, 0, 0, 0, 0, 0)
vec

# replacing on value
vec[1] <- 1000
vec

# replacing multiple values
vec[c(1, 3, 5)] <- c(1, 1, 1)
vec

# crating new values
vec[7] <- 0
vec
```


### Adding nw variabeles to data set
```{r}
deck$new <- 1:52
head(deck)

# Can also remove variables
deck$new <- NULL
head(deck)
```

### Wokring with the deck - Card game War, relpacing aces valu of 1 to 14
```{r}
# singling out value of the aces
deck[c(13, 26, 39, 52), ]

# can just get the values
deck$value[c(13, 26, 39, 52)]

# replacing ace value
deck$value[c(13, 26, 39, 52)] <- 14
head(deck, 13)
```


## Logical Subsetting
```{r}
# Logical Opereators Examples:
1 > 2
1 > c(0, 1, 2)
c(1, 2, 3) == c(3, 2, 1)

# %in% operator
1 %in% c(3, 4, 5)
c(1, 2, 3, 4) %in% c(3, 4, 5)
```
### Exercise: Extract the face column o fdeck and test whether each valu is equal to ace and count how many cards aree equal to ace. 
```{r}
sum(deck$face == "ace")
```

### Exeercise: Lets put logical subsetting to use with a new gam: hearts. In haearts everey card has a valu of zero:
```{r}
deck4 <- deck
deck4$value <- 0

# Assign a value of 1 to every card in dck4 that has a suit of harts
deck4$value[deck4$suit == "hearts"] <- 1
deck4$value[deck4$suit == "hearts"]
```

## Boolan Operators: We can use boolean opeerators to find the queen of spades and assign it the value of 13 for the game of hearts
```{r}
queenSpades <- deck4$face == "queen" & deck4$suit == "spades"
deck4$value[queenSpades] <- 13
deck4[queenSpades, ]
```

## One last game: Black Jack:
```{r}
deck5 <- deck

facecards <- deck5$face %in% c("king", "queen", "jack")
deck5[facecards, ]

deck5$value[facecards] <- 10
head(deck5, 13)
```


## Missing Informtaion:
```{r}
1 + NA

NA == 1
```

### na.rm
```{r}
c(NA, 1:50)

# cant get mean regularly
mean(c(NA, 1:50))

# Use special parameters
mean(c(NA, 1:50), na.rm = TRUE)
```
### is.na
```{r}
# May want to identify if a value is NA
NA == NA

# Regular comparison oerators dont work so use is.na
is.na(NA)
```
### Last thing we will finish the blckjack deck by making the values NA since we dont know the inal value of the ace
```{r}
deck5$value[deck5$face == "ace"] <- NA
head(deck5, 13)
```

